---
title: "BlindingProtocol"
author: "Alexandra Sarafoglou"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# General Info

This script reads in two data files (corresponding to the two session data). It blinds the original data and plots the individual response pattern for the blinded data. It outputs the following files:

- 'blinding_table.csv': This file is for JH only is needed to unblind the data later on
- 'data_blinded.csv': This file is the blinded data which can be shared with AS and HG (not necessary to share)
- 'exclusion_table.csv': This file encodes which participants and quantifier exclusions should be excluded from the data and should be shared with AS and HG
- html output: The html output contains the blinded data pattern and should be shared with AS and HG

```{r load-data}
rm(list=ls())
# shuffle sessions: "../PathToSession1", "../PathToSession2"
session_datafiles   <- c("../data/exp1-replication-trials.csv",
                         "../data/exp1-replication-trials.csv")
session_datafiles   <- sample(session_datafiles)

# read in data
dat_orig_list <- list(dat_orig1 = read.csv(session_datafiles[1]),
                      dat_orig2 = read.csv(session_datafiles[2]))
quant     <- c('Few', 'Fewer than half', 'Many', 'More than half', 'Most')
```

- We flipped the true and false responses for few and fewer than half
- We removed the quantifiers "all", "none", and "some" from the analysis

```{r preprocessing, echo = FALSE}
# preprocess for each element in list
for(i in 1:2) {
  
  dat_orig_list[[i]]$quant <- gsub('\"', '', dat_orig_list[[i]]$quant)
  dat_orig_list[[i]]       <- subset(dat_orig_list[[i]], quant %in% c('Few', 'Fewer than half', 'Many',
                                        'More than half', 'Most'))
  quant_many <- c('Many')
  quant_pos  <- c('More than half', 'Most')
  quant_neg  <- c('Few', 'Fewer than half')
  dat_orig_list[[i]]$qq     <- factor(dat_orig_list[[i]]$quant)
  
  dat_orig_list[[i]]$resp     <- as.numeric(factor(dat_orig_list[[i]]$response)) - 1
  dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_pos,]$resp <- ifelse(dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_pos,]$response=="true",1 , 0)
  dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_neg,]$resp <- ifelse(dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_neg,]$response=="true",0 , 1)
  # double flip 'many'
  dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_many,]$resp    <- ifelse(dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_many,]$response=="true", 0 , 1)
  dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_many,]$percent <- 100 - dat_orig_list[[i]][dat_orig_list[[i]]$quant %in% quant_many,]$percent
  
}

```

- Preregistered: quantifier and session cannot be identified
- Data for each participant should stay the same (to identify participants who need to be excluded)
- Quantifier labels should be shuffled within each participant
- Worker ids stay intact to allow for matching between session 1 and session 2 
- also look at histograms of reaction times for trial exclusion

- Combine session 1 and session 2 data

```{r combine-datasets}
# combine datasets; distinguish between quantifiers in session 1 and 2
for(i in 1:2) {
  
  dat_orig_list[[i]]$quant <- paste0(dat_orig_list[[i]]$quant, i)
  dat_orig_list[[i]]$qq    <- paste0(dat_orig_list[[i]]$qq, i)
  dat_orig_list[[i]]$qq    <- factor(dat_orig_list[[i]]$qq)
  
}

dat_orig <- rbind(dat_orig_list[[1]], dat_orig_list[[2]])
```

```{r blind-condition-labels, echo = FALSE}
# shuffle the labels for each worker id differently
quantifiers          <- unique(dat_orig$quant)
nr_quants            <- length(quantifiers)
quant_colors         <- RColorBrewer::brewer.pal(name='Set3', n=10)
blinded_quant_labels <- paste0('Quantifier', 1:nr_quants)
subjects             <- unique(dat_orig$workerid)
nr_subjs             <- length(unique(dat_orig$workerid))

dat <- data.frame(workerid=NA, A=NA, B=NA, percent=NA, read_and_decide_time=NA,
                  quant=NA, read_time_one=NA, resp=NA)
blinding_table <- as.data.frame(matrix(NA, 
                                       ncol=(1+nr_quants), 
                                       dimnames=list(1, c('workerid', quantifiers))))

for(i in 1:nr_subjs){
  
  # shuffle labels for each person separately
  new_quant_ids      <- sample(1:nr_quants)
  new_quant_labels  <- paste0('Quantifier', new_quant_ids)
  blinding_table[i,] <- c(subjects[i]     , new_quant_ids)
  
  subset             <- dat_orig[dat_orig$workerid == subjects[i], ]
  subset$quant       <- do.call(dplyr::recode, c(list(subset$quant), setNames(new_quant_labels, quantifiers)))

  
  dat      <- rbind(dat, subset[, colnames(subset) %in% colnames(dat)])
  
}

write.csv(blinding_table, file='blinding_table.csv', row.names = FALSE)

# remove first row
dat    <- dat[-1,]
dat$qq <- factor(dat$quant)

write.csv(dat, file='data_blinded.csv', row.names = FALSE)
```

Ensure that the data are effectively blinded for both sessions

```{r visualize-blinded-data, eval = FALSE, echo = FALSE}
par(mfrow=c(1, 2))
# original data
prop <- tapply(dat_orig$resp, list(dat_orig$percent, dat_orig$qq), mean, na.rm = T)
matplot(as.numeric(rownames(prop)), prop
        , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
        , xlab = "Percent", ylab = "Proportion 'true' responses"
        , frame.plot = F)
legend(75, .7, legend = levels(dat_orig$qq), fill = quant_colors, bg = "white", box.col = "white")

# blinded data
qcols <- RColorBrewer::brewer.pal(max(length(levels(dat$qq)), 3), "Dark2")[1:length(levels(dat$qq))]

prop <- tapply(dat$resp, list(dat$percent, dat$qq), mean, na.rm = T)
matplot(as.numeric(rownames(prop)), prop
        , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
        , xlab = "Percent", ylab = "Proportion 'true' responses"
        , frame.plot = F)
legend(75, .7, legend = levels(dat$qq), fill = quant_colors, bg = "white", box.col = "white")
```

```{r export-exclusion-file, echo = FALSE}
exclusion_table <- data.frame(worker_id  = subjects,
                              exclude_all = FALSE,
                              Quantifier1 = FALSE,
                              Quantifier2 = FALSE,
                              Quantifier3 = FALSE,
                              Quantifier4 = FALSE,
                              Quantifier5 = FALSE,
                              Quantifier6 = FALSE,
                              Quantifier7 = FALSE,
                              Quantifier8 = FALSE,
                              Quantifier9 = FALSE,
                              Quantifier10 = FALSE)
# add some exclusions to try out the code
# exclusion_table[c(1, 3, 5, 6), 'exclude_all'] <- TRUE
# exclusion_table[2, sample(paste0('Quantifier', 1:10), 3)] <- TRUE
# exclusion_table[4, sample(paste0('Quantifier', 1:10), 2)] <- TRUE
write.csv(exclusion_table, file='exclusion_table.csv', row.names = FALSE)
```

The visualizations is what will be given to AS and HG.
Based on these visualizations we will exclude additional participants.
Note: if someone is excluded in one timepoint, exclude them all together

```{r visualize-participant-responses, echo = FALSE}
n_observed <- 6

par(mfrow=c(2, 3))

# compute observed proportions
percent_cat       <- cut(dat$percent, breaks=seq(0, 100, length.out=n_observed+1))
prop_observed_ind <- tapply(dat$resp, list(percent_cat, dat$workerid, dat$qq), mean, na.rm = T)
props_lower       <- c(seq(0, 100, length.out=n_observed+1) + 1)[-(n_observed+1)]
props_upper       <- c(seq(0, 100, length.out=n_observed+1))[-1]
cperc_observed    <- (props_upper - ((props_upper - props_lower) / 2))
cperc_observed    <- (cperc_observed - 50) / 100
quant_labels      <- dimnames(prop_observed_ind)[[3]]

for(i in 1:nr_subjs){
  
  for(d in 1:nr_quants){

    plot(cperc_observed, prop_observed_ind[, toString(subjects[i]), blinded_quant_labels[d]],
         type = 'l', las = 1, lwd = 2, bty = 'n',
         xaxt = 'n', yaxt = 'n',
         main = paste0('Subject ', subjects[i], ', ', blinded_quant_labels[d]),
         col  = quant_colors[d], cex.lab=1.2,
         xlab = 'Presented Proportion', 
         ylab = 'Proportion of "true" responses',
         xlim = c(-0.5, 0.5),
         ylim = c(-0.1, 1.1))
    axis(side =2, at = c(0, 0.5, 1), labels = c(0, 0.5, 1), las = 1)
    axis(side =1, at = seq(-0.5, 0.5, by = 0.2), labels = seq(0, 100, by = 20))
    
    # add points to the plot
    observed_responses <- dat[dat$workerid==subjects[i] & 
                              dat$qq == blinded_quant_labels[d], c('resp', 'percent')]
    observed_responses$percent <- (observed_responses$percent - 50) / 100
    observed_responses$resp    <- ifelse(observed_responses$resp == 1, 1.1, -0.1)
    points(observed_responses$percent, observed_responses$resp, col=quant_colors[d])
    
  }
  
  plot.new()
  plot.new()
  
}
```

<!-- Send html output of this script to AS and HG. If they send back their exclusion table, run  -->
<!-- "blinding_RTexclusions" next -->


