---
title: "BlindingProtocol"
author: "Alexandra Sarafoglou"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r load-data}
rm(list=ls())
source('../helper_functions/coorporate_design.R')
dat_orig <- read.csv("../data/exp1-replication-trials.csv")
quant    <- c('Few', 'Fewer than half', 'Many', 'More than half', 'Most')
```

- We flipped the true and false responses for few and fewer than half
- We removed the quantifiers "all", "none", and "some" from the analysis

```{r preprocessing, echo = FALSE}
dat_orig$quant <- gsub('\"', '', dat_orig$quant)
dat_orig       <- subset(dat_orig, quant %in% c('Few', 'Fewer than half', 'Many',
                                      'More than half', 'Most'))
quant_many <- c('Many')
quant_pos  <- c('More than half', 'Most')
quant_neg  <- c('Few', 'Fewer than half')
dat_orig$qq     <- factor(dat_orig$quant)

dat_orig$resp     <- as.numeric(factor(dat_orig$response)) - 1
dat_orig[dat_orig$quant %in% quant_pos,]$resp <- ifelse(dat_orig[dat_orig$quant %in% quant_pos,]$response=="true",1 , 0)
dat_orig[dat_orig$quant %in% quant_neg,]$resp <- ifelse(dat_orig[dat_orig$quant %in% quant_neg,]$response=="true",0 , 1)
# double flip for many
dat_orig[dat_orig$quant %in% quant_many,]$resp    <- ifelse(dat_orig[dat_orig$quant %in% quant_many,]$response=="true", 0 , 1)
dat_orig[dat_orig$quant %in% quant_many,]$percent <- 100 - dat_orig[dat_orig$quant %in% quant_many,]$percent
```

- Preregistered: quantifier and session cannot be identified
- Data for each participant should stay the same (to identify participants who need to be excluded)
- Quantifier labels should be shuffled within each participant
- Worker ids stay intact to allow for matching between session 1 and session 2 

Needs to be added: blinding for session 2

```{r blind-condition-labels, echo = FALSE}
# shuffle the labels for each worker id differently
quantifiers          <- unique(dat_orig$quant)
blinded_quant_labels <- paste('Quantifier', 1:5)
subjects             <- unique(dat_orig$workerid)
nsubjs               <- length(unique(dat_orig$workerid))

dat <- data.frame(workerid=NA, A=NA, B=NA, percent=NA, read_and_decide_time=NA,
                  quant=NA, read_time_one=NA, resp=NA)
blinding_table <- as.data.frame(matrix(NA, ncol=6, 
                                       dimnames=list(1, c('workerid', quantifiers))))

for(i in 1:nsubjs){
  
  # shuffle labels for each person separately
  new_quant_ids      <- sample(1:5)
  new_quant_labels   <- paste('Quantifier', new_quant_ids)
  blinding_table[i,] <- c(subjects[i]     , new_quant_ids)
  
  subset             <- dat_orig[dat_orig$workerid == subjects[i], ]
  subset$quant       <- do.call(dplyr::recode, c(list(subset$quant), setNames(new_quant_labels, quantifiers)))

  
  dat      <- rbind(dat, subset[, colnames(subset) %in% colnames(dat)])
  
}

write.csv(blinding_table, file='blinding_table.csv', row.names = FALSE)

# remove first row
dat    <- dat[-1,]
dat$qq <- factor(dat$quant)
```

Ensure that the data are effectively blinded for both sessions

```{r visualize-blinded-data-session-1, eval = FALSE, echo = FALSE}
par(mfrow=c(1, 2))
# original data
prop <- tapply(dat_orig$resp, list(dat_orig$percent, dat_orig$qq), mean, na.rm = T)
matplot(as.numeric(rownames(prop)), prop
        , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
        , xlab = "Percent", ylab = "Proportion 'true' responses"
        , frame.plot = F)
legend(75, .7, legend = levels(dat_orig$qq), fill = quant_colors, bg = "white", box.col = "white")

# blinded data
qcols <- RColorBrewer::brewer.pal(max(length(levels(dat$qq)), 3), "Dark2")[1:length(levels(dat$qq))]

prop <- tapply(dat$resp, list(dat$percent, dat$qq), mean, na.rm = T)
matplot(as.numeric(rownames(prop)), prop
        , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
        , xlab = "Percent", ylab = "Proportion 'true' responses"
        , frame.plot = F)
legend(75, .7, legend = levels(dat$qq), fill = quant_colors, bg = "white", box.col = "white")
```

```{r visualize-blinded-data-session-2, eval = FALSE, echo = FALSE}
### Repeat the same with session 2 data! ###

# par(mfrow=c(1, 2))
# # original data
# prop <- tapply(dat_orig$resp, list(dat_orig$percent, dat_orig$qq), mean, na.rm = T)
# matplot(as.numeric(rownames(prop)), prop
#         , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
#         , xlab = "Percent", ylab = "Proportion 'true' responses"
#         , frame.plot = F)
# legend(75, .7, legend = levels(dat_orig$qq), fill = quant_colors, bg = "white", box.col = "white")
# 
# # blinded data
# qcols <- RColorBrewer::brewer.pal(max(length(levels(dat$qq)), 3), "Dark2")[1:length(levels(dat$qq))]
# 
# prop <- tapply(dat$resp, list(dat$percent, dat$qq), mean, na.rm = T)
# matplot(as.numeric(rownames(prop)), prop
#         , pch = 19, col = quant_colors, type='l', lwd = 2, lty=1
#         , xlab = "Percent", ylab = "Proportion 'true' responses"
#         , frame.plot = F)
# legend(75, .7, legend = levels(dat$qq), fill = quant_colors, bg = "white", box.col = "white")
```

```{r export-exclusion-file, echo = FALSE}
exclusion_table <- data.frame(worker_id  = subjects,
                              exclude_all = FALSE,
                              exclude_quantifier_1 = FALSE,
                              exclude_quantifier_2 = FALSE,
                              exclude_quantifier_3 = FALSE,
                              exclude_quantifier_4 = FALSE,
                              exclude_quantifier_5 = FALSE)
write.csv(exclusion_table, file='exclusion_table.csv', row.names = FALSE)
```

The visualizations is what will be given to AS and HG.
Based on these visualizations we will exclude additional participants.
Note: if someone is excluded in one timepoint, exclude them all together

```{r visualize-participants-session-1, echo = FALSE}
n_observed <- 6

par(mfrow=c(2, 3))

# compute observed proportions
percent_cat       <- cut(dat$percent, breaks=seq(0, 100, length.out=n_observed+1))
prop_observed_ind <- tapply(dat$resp, list(percent_cat, dat$workerid, dat$qq), mean, na.rm = T)
props_lower       <- c(seq(0, 100, length.out=n_observed+1) + 1)[-(n_observed+1)]
props_upper       <- c(seq(0, 100, length.out=n_observed+1))[-1]
cperc_observed    <- (props_upper - ((props_upper - props_lower) / 2))
cperc_observed    <- (cperc_observed - 50) / 100
quant_labels      <- dimnames(prop_observed_ind)[[3]]

for(i in 1:nsubjs){
  
  for(d in 1:5){

    plot(cperc_observed, prop_observed_ind[, i, d],
         type = 'l', las = 1, lwd = 2, bty = 'n',
         xaxt = 'n', yaxt = 'n',
         main = paste0('Subject ', subjects[i], ', ', blinded_quant_labels[d]),
         col  = quant_colors[d], cex.lab=1.2,
         xlab = 'Presented Proportion', 
         ylab = 'Proportion of "true" responses',
         xlim = c(-0.5, 0.5),
         ylim = c(-0.1, 1.1))
    axis(side =2, at = c(0, 0.5, 1), labels = c(0, 0.5, 1), las = 1)
    axis(side =1, at = seq(-0.5, 0.5, by = 0.2), labels = seq(0, 100, by = 20))
    
    # add points to the plot
    observed_responses <- dat[dat$workerid==subjects[i] & 
                              dat$qq == blinded_quant_labels[d], c('resp', 'percent')]
    observed_responses$percent <- (observed_responses$percent - 50) / 100
    observed_responses$resp    <- ifelse(observed_responses$resp == 1, 1.1, -0.1)
    points(observed_responses$percent, observed_responses$resp, col=quant_colors[d])
    
  }
  
  plot.new()
  
}

```

```{r visualize-participants-session-2, echo = FALSE}
### Repeat the same with session 2 data! ###

# n_observed <- 6
# 
# par(mfrow=c(2, 3))
# 
# # compute observed proportions
# percent_cat       <- cut(dat$percent, breaks=seq(0, 100, length.out=n_observed+1))
# prop_observed_ind <- tapply(dat$resp, list(percent_cat, dat$workerid, dat$qq), mean, na.rm = T)
# props_lower       <- c(seq(0, 100, length.out=n_observed+1) + 1)[-(n_observed+1)]
# props_upper       <- c(seq(0, 100, length.out=n_observed+1))[-1]
# cperc_observed    <- (props_upper - ((props_upper - props_lower) / 2))
# cperc_observed    <- (cperc_observed - 50) / 100
# quant_labels      <- dimnames(prop_observed_ind)[[3]]
# 
# for(i in 1:nsubjs){
#   
#   for(d in 1:5){
# 
#     plot(cperc_observed, prop_observed_ind[, i, d],
#          type = 'l', las = 1, lwd = 2, bty = 'n',
#          xaxt = 'n', yaxt = 'n',
#          main = paste0('Subject ', subjects[i], ', ', blinded_quant_labels[d]),
#          col  = quant_colors[d], cex.lab=1.2,
#          xlab = 'Presented Proportion', 
#          ylab = 'Proportion of "true" responses',
#          xlim = c(-0.5, 0.5),
#          ylim = c(-0.1, 1.1))
#     axis(side =2, at = c(0, 0.5, 1), labels = c(0, 0.5, 1), las = 1)
#     axis(side =1, at = seq(-0.5, 0.5, by = 0.2), labels = seq(0, 100, by = 20))
#     
#     # add points to the plot
#     observed_responses <- dat[dat$workerid==subjects[i] & 
#                               dat$qq == blinded_quant_labels[d], c('resp', 'percent')]
#     observed_responses$percent <- (observed_responses$percent - 50) / 100
#     observed_responses$resp    <- ifelse(observed_responses$resp == 1, 1.1, -0.1)
#     points(observed_responses$percent, observed_responses$resp, col=quant_colors[d])
#     
#   }
#   
#   plot.new()
#   
# }

```


### Exclusion criteria - Participants

- We excluded participants who had 50% or more reaction times faster than 300ms
- We excluded participants who failed to obey the monotonicity of quantifiers (by visual inspection) 

### Exclusion criteria - Trials

- We excluded trials with response times shorter than 300ms 
- We excluded trials longer than 2500ms


### Exclude Participants and Trials based on visual inspection

```{r blinded-exclusion, eval = FALSE}
blinding_table <- read.csv('blinding_table.csv' , header = TRUE)
exclusions     <- read.csv('exclusion_table.csv', header = TRUE)
# if exclusion_all = TRUE, fill in remainder of the columns with TRUE
exclusions[exclusions$exclude_all==TRUE, grepl('exclude_quant', colnames(exclusions))] <- TRUE

# map blinded data to original data

for(i in 1:nsubjs){
  
   exclusions[i, 3:7] <- exclusions[i, 3:7][order(unlist(blinding_table[i, 2:6]))]
  
}

```

