---
title: "BlindingProtocol"
author: "Alexandra Sarafoglou"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# General Info

This script reads in the blinded data and the exclusion table. It excludes participants based on the exclusions declared in the exclusion table. Make sure AS and HG send you their exclusion table, before running this script. Then, for the remaining participants and quantifier conditions, it plots fast and slow responses. It also excludes fast and slow responses. Make sure AS and HG send you their cutoff criteria for fast and slow responses. This script outputs the following files:

- html output: The html output contains the blinded reponse time distributions and should be shared with AS and HG before continuing to preprocess the data.
- 'data_blinded_preprocessed.csv': This file is preprocessed blinded data (with participant and response exclusions) which can be shared with AS and HG for further analysis


# Exclude Participants and Quantifiers based on visual inspection

Make sure AS and HG send you their exclusion table, before running this script.

```{r load-data, echo = FALSE}
rm(list=ls())
dat            <- read.csv('data_blinded.csv' , header = TRUE)
# one exclusion table with joint decision from AS and HG
exclusions     <- read.csv('exclusion_table.csv', header = TRUE) # try csv2 if separator is ";"
quantifiers    <- unique(dat$quant)
```

```{r blinded-exclusion, echo = FALSE}
subjects  <- unique(dat$workerid)
total_N   <- nrow(exclusions)
nr_quants <- length(unique(dat$quant))

nr_quants_excluded <- 0
dat$keep           <- TRUE

for(i in 1:total_N){
  
  # check whether the participant needs to be excluded
  exclude_participant <- exclusions[exclusions$worker_id == subjects[i], 'exclude_all']
  
  if(exclude_participant){
    
    dat[dat$workerid == subjects[i], 'keep'] <- FALSE
    
  } else {
    
    exclude_any_quants <- any(exclusions[exclusions$worker_id == subjects[i], 
                                         quantifiers])
    
    if(exclude_any_quants) {
      
      quants_to_exclude  <- quantifiers[unlist(exclusions[exclusions$worker_id == subjects[i],
                                                          quantifiers])]
      
      for(j in seq_along(quants_to_exclude)) {
        
        nr_quants_excluded <- nr_quants_excluded + 1
        dat[dat$workerid == subjects[i] & dat$quant == quants_to_exclude[j], ] <- FALSE
        
      }
    } 
  }
}

dat <- dat[dat$keep, ]
rm(subjects, nr_quants)
```

```{r, echo = FALSE}
# new variables based on excluded data
subjects    <- unique(dat$workerid)
nr_subjs    <- length(unique(dat$workerid))
nr_excluded <- total_N - nr_subjs
nr_quants   <- length(unique(dat$quant))
```

We excluded `r nr_excluded` out of  `r total_N` participants, which corresponds to `r round((nr_excluded/total_N) * 100, 2)` percent of the sample. From the remaining sample, we excluded `r nr_quants_excluded` quantifier conditions from the data. 

## Histogram of fast response times

```{r visualize-fast-responses-all, echo = FALSE}
# visualize everything below 1000ms
hist(dat[dat$read_and_decide_time < 1000, 'read_and_decide_time'], 
     main   = 'All participants',
     las    = 1,
     xlab   = 'Fast responses (in ms)',
     border = 'white',
     col    = '#80B1D3')
abline(v=c(150, 200, 250, 300),
       col=c('#A6761D', '#1B9E77', '#7570B3', 'red'),
       lwd = 2,
       lty = c(3, 3, 3, 1))
legend('topright',
       legend = c(150, 200, 250, 300),
       col=c('#A6761D', '#1B9E77', '#7570B3', 'red'),
       lty = c(3, 3, 3, 1),
       lwd = 2
       )
```

```{r visualize-fast-responses, echo = FALSE}
# visualize everything below 1000ms
par(mfrow=c(2, 3))
for(i in 1:nr_subjs){
  
  hist(dat[dat$workerid == subjects[i] & dat$read_and_decide_time < 1000, 'read_and_decide_time'], 
       main   = paste('Subject', subjects[i]),
       las    = 1,
       xlab   = 'Read and decide time (in ms)',
       xlim   = c(0, 1000),
       border = 'white',
       col    = '#80B1D3')
  abline(v=c(150, 200, 250, 300),
         col=c('#A6761D', '#1B9E77', '#7570B3', 'red'),
         lwd = 2,
         lty = c(3, 3, 3, 1))

}
```


## Histogram of slow response times

```{r visualize-slow-responses-all, echo = FALSE}
hist(dat[dat$read_and_decide_time < 5000, 'read_and_decide_time'], 
     main   = 'All participants',
     las    = 1,
     xlab   = 'Slow responses (in ms)',
     border = 'white',
     col    = '#80B1D3')
abline(v=c(2500, 3000, 3500, 4000),
       col=c('red', '#A6761D', '#1B9E77', '#7570B3'),
       lwd = 2,
       lty = c(1, 3, 3, 3))
legend('topright',
       legend = c(2500, 3000, 3500, 4000),
       col=c('red', '#A6761D', '#1B9E77', '#7570B3'),
       lty = c(1, 3, 3, 3),
       lwd = 2
       )
```

```{r visualize-slow-responses, echo = FALSE}
par(mfrow=c(2, 3))
for(i in 1:nr_subjs){
  
  hist(dat[dat$workerid == subjects[i], 'read_and_decide_time'], 
       main   = paste('Subject', subjects[i]),
       las    = 1,
       xlab   = 'Read and decide time (in ms)',
       border = 'white',
       col    = '#80B1D3')
  abline(v=c(2500, 3000, 3500, 4000),
         col=c('red', '#A6761D', '#1B9E77', '#7570B3'),
         lwd = 2,
         lty = c(1, 3, 3, 3))

}
```

## Exclude fast and slow responses

Make sure AS and HG send you their cutoff criteria for fast and slow responses.

```{r exlcude-responses, eval = FALSE, echo = FALSE}
cutoff_fast <- 300   # based on decision from AS and HG
cutoff_slow <- 2500  # based on decision from AS and HG

dat$fast_response <- FALSE
dat$slow_response <- FALSE

dat[dat$read_and_decide_time < cutoff_fast, 'fast_response'] <- TRUE
dat[dat$read_and_decide_time > cutoff_slow, 'slow_response'] <- TRUE
```

From all `r nrow(dat)` trials, we excluded `r sum(dat$fast_response)` fast responses from the data, which corresponds to `r round((sum(dat$fast_response)/nrow(dat)) * 100, 2)` percent of all trials. In addition, we excluded `r sum(dat$slow_response)` slow responses from the data, which corresponds to `r round((sum(dat$slow_response)/nrow(dat)) * 100, 2)` percent of all trials.

```{r export-preprocessed-data, eval = FALSE, echo = FALSE}
dat <- dat[dat$read_and_decide_time > cutoff_fast &  
           dat$read_and_decide_time < cutoff_slow,]

write.csv(dat, file='data_blinded_preprocessed.csv', row.names = FALSE)
```



